---
title: "p8105_hw3_yx2640"
author: "Elaine Xu"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

There are 134 aisles, most items are from fresh vegetables.
```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


### Problem 2

* Load, tidy, and otherwise wrangle the data.
```{r}
accel =
  read_csv(
		"./data/accel_data.csv"
		) %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_ct",
    values_to = "minute") %>%
  mutate(
    day_num = recode(day, "Sunday" = "7", "Saturday" = "6", "Friday" = "5", "Monday" = "1", "Tuesday" = "2", "Wednesday" = "3", "Thursday" = "4")) %>%
  arrange(week, day_num) %>%
  select(-day_num) %>%
  relocate(minute, .before = "activity_ct") %>%
  mutate(
    weekday_vs_weekend = case_when(
    day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
    day %in% c("Saturday","Sunday") ~ "weekend")) %>%
  mutate(
    day = as.factor(day))
```

This data collected five weeks of accelerometer data on a 63 year-old male with BMI 25. After tidy the original data, the resulting dataset has `r nrow(accel)` observations and `r ncol(accel)` variables. Day of the week is the only factor class variable. Other then `day`, variables `week`, `day_id` and `minute` are class `numeric`, and variables `activity_ct` and `weekday_vs_weekend` are class `numeric`. Variable `minute` collected 1440 minutes of each day in the `day` variable over five weeks.


*  Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals.
```{r}
total_active =
  accel %>%
	group_by(week, day) %>%
	count(sum(minute)) %>%
  mutate(
    day = as.character(day),
    day_num = recode(day, "Sunday" = "7", "Saturday" = "6", "Friday" = "5", "Monday" = "1", "Tuesday" = "2", "Wednesday" = "3", "Thursday" = "4")) %>%
  arrange(week, day_num) %>%
  select(-day_num, -n) %>%
  janitor::clean_names() %>%
  pivot_wider(names_from = day, values_from = sum_minute)

knitr::kable(total_active)
```
